Lay Perceptions of Crowd-Scientific Findings: The Risks of Variability and Lack of Consensus

### 2021-10-07 manuscript/manuscript.Rmd
manuscript: fixed typo, talk: started table many-analysts studies

Crowd science is defined as "the organization of scientific research in open collaborative projects"  [@franzoni2014, p. 1]. Crowd-scientific projects vary in their degree of selectivity: some are open to the general public (such projects are often referred to as "citizen science"), while others require the expertise of specialists [e.g., scientists\; @uhlmann2019, p. 714]. Here, we focus on crowdsourced data analysis, also known as the 'many analysts' or 'multi-analyst' approach: giving the same dataset to different teams of scientists, who independently analyze it to answer the same research question and/or estimate a parameter of interest.  

@@

According to previous work, crowdsourced data analysis should increase the credibility of scientific findings; Table 1 provides an overview of the proposed confidence- and credibility-related benefits of the multi-analyst approach. Does crowd science meet its promise -- to improve the credibility of scientific research -- in reality? Normatively, we expect laypeople who observe several scientists independently come to the same qualitative conclusions -- compared to a single research team that comes to a single, joint conclusion -- to be more swayed by the findings, as it seems to reflect consensus between various approaches. After all, it's a hallmark of rigorous research to triangulate various methodological approaches and provide converging evidence [@jick1979; @turner2017], and to provide a "sensitivity analysis" or "robustness checks" to examine whether and how the findings change as a result of alternative analytic specifications [@munoz2018; @steegen2016]. As argued in Aczel et al. [-@aczel2020, p. 562], "The main argument for the importance of performing robustness checks over reasonable variations in modelling choices is to increase confidence in the obtained results: ideally, results should be reasonably unaffected by a researcher's idiosyncratic choice (…) when reasonable alternatives exist." The same argument has been made for triangulation, which, according to Jick [-@jick1979, p. 608], "allows researchers to be more confident of their results." We would similarly expect readers of multi-analyst reports that yield consistent findings to be more confident in the results. Although lay observers, in such cases, may be less certain that the precise aggregate parameter estimate is exactly correct, they should be more likely to positively update their beliefs about the reported phenomenon, and less likely to assume that the overall estimate stems from bias or error. Thus, when the results generated by independent analysts are largely consistent, we expect an increase in the sway of scientific findings. However, when laypeople observe several scientists independently come to differing qualitative conclusions, we expect the multi-analyst method to backfire; when results across many analysts vary widely and lack consensus in their qualitative conclusions (which, arguably, often reflects the reality of large-scale science collaborations), we expect a decrease in the sway of scientific findings.  

>>>>

Crowd science is defined as "the organization of scientific research in open and collaborative projects"  [@franzoni2014, p. 1]. Crowd-scientific projects vary in their degree of selectivity: some are open to the general public (such projects are often referred to as "citizen science"), while others require the expertise of specialists [e.g., scientists\; @uhlmann2019, p. 714]. Here, we focus on crowdsourced data analysis, also known as the 'many analysts' or 'multi-analyst' approach: giving the same dataset to different teams of scientists, who independently analyze it to answer the same research question and/or estimate a parameter of interest.  

@@

According to previous work, crowdsourced data analysis should increase the credibility of scientific findings; Table 1 provides an overview of the proposed confidence- and credibility-related benefits of the multi-analyst approach. Does crowd science meet its promise -- to improve the credibility of scientific research -- in reality? Normatively, we expect laypeople who observe several scientists independently come to the same qualitative conclusions -- compared to a single research team that comes to a single, joint conclusion -- to be more swayed by the findings, as it seems to reflect consensus between various approaches. After all, it's a hallmark of rigorous research to triangulate various methodological approaches and provide converging evidence [@jick1979; @turner2017], and to provide a "sensitivity analysis" or "robustness checks" to examine whether and how the findings change as a result of alternative analytic specifications [@munoz2018; @steegen2016]. As argued in Aczel et al. [-@aczel2020, p. 562], "The main argument for the importance of performing robustness checks over reasonable variations in modelling choices is to increase confidence in the obtained results: ideally, results should be reasonably unaffected by a researcher's idiosyncratic choice (…) when reasonable alternatives exist." The same argument has been made for triangulation, which, according to Jick [-@jick1979, p. 608], "allows researchers to be more confident of their results."  
We similarly expect readers of multi-analyst reports that yield consistent findings to be more confident in the results. Although lay observers, in such cases, may be less certain that the precise aggregate parameter estimate is exactly correct, they should be more likely to positively update their beliefs about the reported phenomenon, and less likely to assume that the overall estimate stems from bias or error. Thus, when the results generated by independent analysts are largely consistent, we expect an increase in the sway of scientific findings. However, when laypeople observe several scientists independently come to differing qualitative conclusions, we expect the multi-analyst method to backfire; when results across many analysts vary widely and lack consensus in their qualitative conclusions (which, arguably, often reflects the reality of large-scale science collaborations), we expect a decrease in the sway of scientific findings.  

### 2021-10-07 talk/index.Rmd
manuscript: fixed typo, talk: started table many-analysts studies

`r icon_style(icons::fontawesome("users"), scale = 2, fill = "#035AA6")` $~~$ **Crowdsourcing Science** 
`r icon_style(icons::fontawesome("chart-bar"), scale = 2, fill = "#035AA6")` $~~~$ **Crowdsourced Data Analysis**  
`r icon_style(icons::fontawesome("question-circle"), scale = 2, fill = "#035AA6")` $~~~$ **Lay Perceptions: Are Crowd-Scientific Findings More Credible?**   
`r icon_style(icons::fontawesome("tools"), scale = 2, fill = "#035AA6")` $~~~$ **Method**
`r icon_style(icons::fontawesome("chart-line"), scale = 2, fill = "#035AA6")` $~~~$ **Results**
`r icon_style(icons::fontawesome("comment"), scale = 2, fill = "#035AA6")` $~~~$ **Discussion**

@@

Highly **inclusive** (sourcing citizens) or highly **selective** (sourcing specialists)

@@

]
---
.center[
# Crowdsourced data analysis

@@

Also known as the **many-analysts** or **multi-analyst** approach

@@

]
---
exclude: true

```{r}
knitr::kable(head(mtcars), format = 'html')
```

---
exclude: true

```{r eval=requireNamespace("DT", quietly=TRUE)}
DT::datatable(head(mtcars), fillContainer = FALSE, options = list(pageLength = 4))
```

---
exclude: true

## Square image

<center><img src="https://octodex.github.com/images/labtocat.png" height="400px" /></center>

.footnote[GitHub Octocat]

>>>>

`r icon_style(icons::fontawesome("users"), scale = 2, fill = "#035AA6")` $~~$ **Crowdsourcing Science** 
`r icon_style(icons::fontawesome("chart-bar"), scale = 2, fill = "#035AA6")` $~~~$ **Crowdsourcing Data Analysis**  
`r icon_style(icons::fontawesome("question-circle"), scale = 2, fill = "#035AA6")` $~~~$ **Lay Perceptions: Are Crowd-Scientific Findings More Credible?**   
`r icon_style(icons::fontawesome("tools"), scale = 2, fill = "#035AA6")` $~~~$ **Method**
`r icon_style(icons::fontawesome("chart-line"), scale = 2, fill = "#035AA6")` $~~~$ **Results**
`r icon_style(icons::fontawesome("comment"), scale = 2, fill = "#035AA6")` $~~~$ **Discussion**

@@

From highly **inclusive** (sourcing citizens) to highly **selective** (sourcing specialists)

@@

In 1833, Denison Olmsted used letter correspondence to recruit citizen scientists to help document a meteor shower. This led to a detailed documentation of the great meteor shower and the birth of the citizen-science movement (highly inclusive, used for data collection.
---
.center[
# Crowdsourced data analysis

@@

`r icon_style(icons::fontawesome("users"), scale = 2, fill = "#035AA6")` $~~$ Also known as the **many-analysts** or **multi-analyst** approach
`r icon_style(icons::fontawesome("chart-line"), scale = 2, fill = "#035AA6")` $~~~~$ On the **rise** in behavioral science

--

```{r, echo = FALSE}
t1 <- data.frame(
  Authors = c(
    "Botvinik-Nezer et al.",
    "Schweinsberg et al.",
    "Silberzahn et al."
  ),
  Year = c(
    2020,
    2018,
    2018
  ),
  Title = c(
    "Variability in the analysis of a single neuroimaging dataset by many teams",
    "Same data, different conclusions: Radical dispersion in empirical results when independent analysts operationalize and test the same hypothesis",
    "Many Analysts, One Data Set: Making Transparent How Variations in Analytic Choices Affect Results"
  )
)
DT::datatable(t1, fillContainer = FALSE, options = list(pageLength = 2))
```

@@
